An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-01-28 draft.tex
Minor changes and reflow

\address[itt]{Department of Graduate Studies, Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico}
\address[granada]{Universidad de Granada, Granada, Spain}

\begin{abstract} 
Splitting a population into multiple instances is a technique that has been used extensively in
recent years to improve the performance of nature-inspired optimization
algorithms. Work on several populations can be done in parallel, and
also asynchronously, which can be leveraged to create scalable
implementations based on, among other methods, distributed, multi-threaded, parallel, and cloud-based computing. 
% I have rephrased all this. Please check it out - JJ
The design of these cloud-native, distributed, multi-population
algorithms is not a trivial task. Coming from monolithic
(single-instance) solutions, adaptations at several levels, from the
algorithmic to the functional, must be made to leverage
the scalability, elasticity, fault-tolerance, reproducibility, and
cost-effectiveness of cloud systems while, at the same time,
conserving the intended functionality. Instead of an evolutive
approach, in this paper we propose to create from scratch a
cloud-native optimization framework that is able to include multiple
algorithms and use few parameters to work. This solution goes beyond
current state of the art via a framework that is able to support
different algorithms at the same time, work asynchronously, and also
be easily deployable to any cloud platform. We evaluate the
performance and scalability of this solution, together with the effect
other design parameters had on the 
performance, in particular, the number and the size of populations
with respect to problem size. The architecture and the implemented platform is an excellent
alternative for running locally or in the cloud, thus proving that cloud-native
bioinspired algorithms perform better in their "natural" environment than other
kinds of algorithms, and setting a new baseline for scaling and
performance in the cloud.
% Still a big longish. We need another iteration to reduce it - JJ
\end{abstract}

>>>>

\address[itt]{Department of Graduate Studies, Instituto Tecnol\'ogico de Tijuana, Tijuana BC, Mexico}
\address[granada]{Department of Computer Architecture and Technology, Universidad de Granada, Granada, Spain}

\begin{abstract} 
Splitting a population into multiple instances is a technique that has
been used extensively in recent years in order to help improve the
performance of nature-inspired optimization algorithms. Work on
several populations can be done in parallel, and also asynchronously,
a fact that can be leveraged to create scalable implementations based
on, among other methods, distributed, multi-threaded, parallel, and
cloud-based computing.  However, the design of these cloud-native,
distributed, multi-population algorithms is not a trivial task. Coming
from monolithic (single-instance) solutions, adaptations at several
levels, from the algorithmic to the functional, must be made to
leverage the scalability, elasticity, fault-tolerance,
reproducibility, and cost-effectiveness of cloud systems while, at the
same time, conserving the intended functionality. Instead of an
evolutive approach, in this paper we propose to create from scratch a
cloud-native optimization framework that is able to include multiple
algorithms and use few parameters to work. This solution goes beyond
current state of the art via a framework that is able to support
different algorithms at the same time, work asynchronously, and also
be easily deployable to any cloud platform. We evaluate the
performance and scalability of this solution, together with the effect
other design parameters had on it; in particular, the number and the
size of populations with respect to problem size. The architecture and
the implemented platform is an excellent alternative for running
locally or in the cloud, thus proving that cloud-native bioinspired
algorithms perform better in their "natural" environment than other
kinds of algorithms, and setting a new baseline for scaling and
performance in the cloud.
\end{abstract}
